---
title: "Analysis of US Mass Incarceration"
author: "Jasmine Wong"
date: '2022-11-21'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
Systemic racism is an issue that has been occurring in the United States since 
the day it was formed. America's mass incarceration of people of color proves 
that the institution continues to uphold systemic racism and inequality. This 
report analyzes the data about incarceration trends in the U.S. and will show 
how the prison system reinforces oppression and inequality. 


## Introduction 


## Data Summary 
``` {r data summary}

```
Note: Summary values are all from the year 2016

The United States' over policing of people of color is shown in the average 
prison rate 

## Growth of the U.S. Prison Population
```{r, fig.width=10, fig.height=6, echo=FALSE}
plot_jail_pop_for_us()
```

## Growth of Prison Population by State 
``` {r, fig.width=10, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE}
ggplotly(plot_jail_pop_by_states(c("WA", "NY", "CA", "FL", "TX")))
```



## Section 5
``` {r, fig.width=11, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}
plot_female_prison_pops()
```


## Section 6
```{r map, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
plot_map()
```



